---
title: "Tidymodels -- Preprocess your data with recipes"
description: |
  A summary of the tidymodels.org webpage
author:
  - name: Alexis Solis 
    url: alexis.solisc@gmail.com
    # affiliation: Spacely Sprockets
    # affiliation_url: https://example.com/spacelysprokets
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

The `recipes` package is useful in modeing for the preprocessing stage. It helps you with _feature extraction_ and _feature creation_. Recipes are useful for:

1. Creating dummy variables.

2. Transforming your data (e.g. changing to logarithmic scale).

3. Feature extraction - extracting key features from raw variables.


```{r}
library(tidymodels)      # for the recipes package, along with the rest of tidymodels
library(tidyverse)

# Helper packages
library(nycflights13)    # for flight data
library(skimr)           # for variable summaries
library(gt)              # for creating tables.
```

## The New York City flight data

First we load the data and make some changes to the variables:

```{r}
set.seed(123)


flight_data <- 
  flights %>%
  mutate(
    
    # Convert the arrival delay to a factor:
    arr_delay = ifelse(arr_delay >= 30, "late", "on_time"),
    arr_delay = factor(arr_delay),
    
    # We will use the date (not date-time) in the recipe below:
    date = as.Date(time_hour)
  ) %>% 
  
  # Include the weather data. This is done with an inner_join:
  inner_join(weather, by = c("origin", "time_hour")) %>% 
  
  # Only retain useful columns for the analysis:
  select(dep_time, flight, origin, dest, air_time, distance,
         carrier, date, arr_delay, time_hour) %>% 
  
  # Exclude missing data:
  drop_na() %>%
  
  # For creating models, it's better to have qualitative columns
  # encoded as factors (not as strings):
  
  mutate_if(is.character, as.factor)
```

We can get the percentage of flights that arrived late:

```{r}
flight_data %>% 
  count(arr_delay) %>% 
  mutate(prop = n/sum(n))
```
We see that 16% of flights were delayed more than 30 minutes. Since we're doing a logistic regression model, it's important that the response variable is coded as a factor (`fct`). We can check this with the `glimpse()` function:

```{r}
flight_data %>% 
  glimpse()
```
We can use the `skim()` function from the `skimr` package to get a summary of our data. Here we use it to explore two columns: `dest` and `carrier`. From the output, we see that `n_unique` has 104 for `dest` and 16 for carrier (i.e. there are 16 different carriers and 104 different destinations in the dataset).

```{r}
flight_data %>% 
  skimr::skim(dest, carrier)
```

Our model will be a simple logistic regression, so we will need to convert some variables to `dummy` variables.

# DATA SPLITTING
